Evolution of Spectral Distributions in Deep-Water Constant Vorticity Flows

### 2020-01-09 Writeup/random_vorticity_revision_1_5_20.tex
All the updates

\section{Introduction}
There is now a wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.   As explained in \cite{janssen}, MI is especially important due to the relative rapidity with which it acts, being orders of magnitude faster than say four-wave resonant interactions which form the traditional backbone of our understanding of how nonlinearity drives changes in statistics of random-wave fields \cite{hasselmann,holthuijsen}.  These results have significant impact on the understanding of deep-water-rogue wave formation among other oceanographic phenomena.

However, much of the understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.  In particular, most understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.  To address this shortcoming, in now seminal papers, \cite{alber} and \cite{alber2} analytically studied families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.  Using the Nonlinear Schr\"{o}dinger equation (NLSE), an analytic criterion determining when MI is either manifested or suppressed depending on this spectral width, now referred to as a Benjamin--Feir index (BFI) was derived.  The results in \cite{alber} were explored numerically in \cite{dysthe2,dysthe3} by examining the mean properties associated with ensembles of initial conditions.

@@

Later work has simulated the impact of wind and damping on the interplay between the BFI and the kurtosis of a wave field; see \cite{slunyaev,eeltink}.  However, the impact of vorticity was not taken into account in these works, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.  Moreover, as shown recently in \cite{curtis8}, linear shear profiles can have strong impacts on the behavior of mean properties of surface wave flows with some shear profiles being able to suppress the Stokes Drift at the surface.  Therefore while a relatively idealized case of shear, the interplay between nonlinearity and constant vorticity can manifest in markedly different physics.  Building then on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, and complementing the results in \cite{touboul}, we numerically explore the statistical properties of random-nonlinear-wave fields moving over deep water and linear shear profiles.  Moving beyond just examining the properties of the NLSE with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation (VDE) derived in \cite{curtis8}.  As shown in \cite{onorato,janssen} and \cite{eeltink}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity interacts with these higher-order nonlinearities.

Using our simulations, starting from normally distributed random-wave fields, we show that the BFI condition with respect to predicting stability is not especially useful, conforming to the results in \cite{janssen}.  While we are able to show that for sufficiently wide enough initial distributions that the temporal modifications to the profile are slight, whether or not this corresponds to an actual bifurcation in stability or simply reflects that for a wide enough profile the sideband growth affiliated with the MI has little impact is not clear.  We note though that classic MI analysis does provide a useful prediction of the time scale over which the most significant dynamics in the statistics of the wave field occur.

@@

The structure of the paper is an explanation of the model and discussion of the BFI in Section 2.  Section 3 presents our numerical results, while Section 4 presents a conclusion.

\section{Modeling and Stability Theory}
We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we suppose the fluid velocity is of the form
\[
{\bf u} = u({\bf x},t)\hat{{\bf x}} + w({\bf x},t)\hat{{\bf z}} = \omega z\hat{{\bf x}} + \nabla \phi,
\]
where $\phi$ is a harmonic function.  We restrict fluid motion to the $(x,z)$-plane, thereby ignoring transverse variations in the $y$ dimension.   Using well established arguments, e.g.~see \cite{ashton}, the dynamics of the fluid are found by solving the free boundary value problem
\begin{align}
\Delta \phi = 0, & & -\infty< z < \eta(x,t), \nonumber \\
\eta_{t} +\left(\omega\eta+\phi_{x}\right)\eta_{x} - \phi_{z} = 0, & & z=\eta(x,t), \label{kine}\\
\phi_{t} + \omega\pd_{x}^{-1}\eta_{t} + \frac{1}{2} \left|{\bf
    u}\right|^{2} + g\eta -
\frac{\sigma}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
& z=\eta(x,t), \label{bernou}\\
\lim_{z\rightarrow -\infty} \phi_{z} = 0, \nonumber
\end{align}
where $\eta$ represents the free surface displacement, and $g$, $\rho$, and $\sigma$ represent the acceleration due to gravity, the fluid density, and the coefficient of surface tension respectively.  The system is non-dimensionalized via the choices 
\begin{align*}
\tilde{x} = \frac{x}{L_{p}}, ~\tilde{z} = \frac{z}{L_{p}}, ~ \tilde{t} = \sqrt{\frac{g}{L_{p}}}~t, ~\omega = \sqrt{\frac{g}{L_{p}}}~\tilde{\omega}, \\
\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL_{p}}~\tilde{\phi} , ~ \epsilon = \frac{a}{L_{p}}.
\end{align*}
Note that we are assuming $\omega = \mathcal{O}(1)$ throughout the remainder of the paper.  In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L_{p}/g}$.  If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and asymptotic analysis would be siginificantly more difficult.  Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.

For a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$, we model a modulated-wave packet via the expansion 
\begin{multline*}
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta(x,t)} + \eta^{\ast}_{1}(\xi,\tau)e^{-i\theta(x,t)}\\
+ \epsilon\left(\eta_{0}(\xi,\tau) + \eta_{2}(\xi,\tau)e^{2i\theta(x,t)} + \eta^{\ast}_{2}(\xi,\tau)e^{-2i\theta(x,t)}\right) + \mathcal{O}\left(\epsilon^{2} \right)
\end{multline*}
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and $\theta(x,t) = k_{0}x + \Omega(k_{0},\omega)t$, where the linear dispersion relationship $\Omega(k_{0},\omega)$ is given by
\[
\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}k_{0}^{2}\right)} \right).
\]
We note that the form for $\phi$ follows directly from this expansion, allowing for everything to be modeled by the terms in the expansion in $\eta$.  Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
\[
\tilde{\sigma} = \frac{\sigma}{\rho g L_{p}^2}.
\]
Throughout the remainder of this paper, we take the positive branch $+$ of the dispersion relationship, thereby ensuring that $\Omega > 0$.  Thus, for $k_{0}>0$, the surface waves to leading order propagate to the left.  Thus, if $\omega>0$, then throughout the bulk of the fluid, the shear current propagates with the surface waves, or is {\it co-propagating}, while for $\omega<0$, we say the shear current is {\it counter-propagating}.

To describe the evolution of the slowly evolving envelope $\eta_{1}(\xi,\tau)$, we use two models.  The first model is the the nonlinear Schr\"{o}dinger equation (NLSE) derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by 
\[
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} = 0, 
\]
where 
\begin{align*}
c_{g} = & \frac{1+3\tilde{\sigma}k_{0}^{2}}{2s\Omega - \omega},\\
\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma})}{2\Omega-s\omega},\\
\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}k_{0}^{2} + 2(\tilde{\sigma}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}k_{0}^{2})+2\omega\Omega)\right)}.
\end{align*}
However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations.  A more robust model, following the work in \cite{dysthe1}, is found by going to the next higher order in $\epsilon$, giving us the equation
\begin{multline}
(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
- i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
+ \epsilon \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1}\\
+ \epsilon\tilde{\alpha}_{2}\eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} +
\epsilon i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left|\eta_{1}\right|^{2}= 0.
\label{dysthe}
\end{multline}
where $\mathcal{H}$ denotes the Hilbert transform so that 
\[
\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
\]
and where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
\[
\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
\]

We call Equation \eqref{dysthe} the Vor-Dysthe equation (VDE).  Details of its derivation can be found in \cite{curtis8}, though we note a difference in this work with that in \cite{curtis8}.  In this work, we found the inclusion of the higher-order dispersion associated with the coefficient $c_{h}$, given by
\[
c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma})/(\omega-2s\Omega) - s\alpha_{d}^{2},
\]
was necessary to remove artificial numerical instabilities.  Similar issues with Dysthe like equations have been noted in \cite{eeltink} among other sources.  Moreover, the coefficients $\tilde{\alpha}_{j}$ are readily derived, but their forms are cumbersome and thus we omit explicitly writing them for the sake of readability.

To generate simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L},
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly and independently chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}(\xi,0)\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, motivated by the work of \cite{alber} and \cite{dysthe2} and more fully explored in \cite{janssen}, we study the stability of the initial wavepacket distributions with regards to the magnitude of the Benjamin--Feir Index (BFI), where the BFI is defined to be  
\[
\mbox{BFI} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\frac{\epsilon_{rms}}{\sigma}.
\]
Note, this definition is identical with the spectral width criterion of \cite{alber} and the BFI condition given in \cite{thomas2012nonlinear}; see also \cite{onorato}.  It was claimed in \cite{alber}, and in some sense verified in \cite{dysthe2}, that wave packets should be stable with respect to MI's for $\mbox{BFI}<1$, which corresponds to the initial spectral width $\sigma>\sigma_{c}$, where 
\[
\sigma_{c} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\epsilon_{rms}.
\]

However, \cite{alber} does not present numerical verification of their results, and \cite{dysthe2} is relatively brief in their verification of the BFI condition.  The work in \cite{janssen} shows that the BFI condition does not yield a clear bifurcation in stability.  Our results likewise support this.  However, also shown in \cite{janssen} and followed up on in several later works; see \cite{onorato,slunyaev,eeltink}, it is shown that the kurtosis of the corresponding distribution of the surface waves, say $\tilde{k}(\tau)$ is a quadratic function of the time evolving $\mbox{BFI}(\tau)$.   We likewise aim to establish a similar relationship, thereby maintaining the utility of the BFI insofar as it then helps predict a key feature of nonlinear-random wave fields. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
To understand the relationship between the vorticity $\omega$ and the initial spectral width $\sigma$ of our initial conditions, we generate ensembles of initial conditions and then run these different initial conditions up to $\tau_{f}$ where $\tau_{f}$ is given by
\[
\tau_{f}=\max\left\{1.5/\alpha_{nl},1\right\},
\]
so that MI has a sufficiently long enough time scale over which to act; see \cite{curtis8} for a derivation of this result.  We choose $L=5\pi$ so that the computational domain is $[-5\pi,5\pi]$.  The number of modes is $K_{T} = 1024$ or $K=512$.  The surface tension is $\tilde{\sigma} = 10^{-5}$. We then compute the averaged spectral density $S(k,\tau;\sigma)$ where 
\[
S\left(k,\tau;\sigma\right) \sim \overline{\left( \left|\hat{\eta}_{1}\left(k,\tau;\sigma \right)\right|^{2}\right)},
\]
which corresponds to the autocorrelation of the surface profile with respect to space.  Note, we normalize $S$ so that it is a probability distribution.  Using $S(k,\tau_{f},\sigma)$, we then plot the time dynamics of the mean $m(\tau)$
\[
m(\tau) = \int_{\mathbb{R}} k S\left(k,\tau;\sigma\right) dk,
\]
standard deviation $\tilde{\sigma}(\tau)$,
\[
\tilde{\sigma}(\tau) = \left(\int_{\mathbb{R}} (k-m(\tau))^{2} S\left(k,\tau;\sigma\right) dk\right)^{1/2},
\]
and the kurtosis $\tilde{k}(\tau)$
\[
\tilde{k}(\tau) = \frac{1}{3\tilde{\sigma}^{4}(\tau)}\int_{\mathbb{R}} (k-m(\tau))^{4} S\left(k,\tau;\sigma\right) dk - 1,
\]
so as to examine the impact of various shear profiles on the statistics of the waves as well as establishing that statistical equilibrium has been achieved by $\tau_{f}$.  Note, we interpret the kurtosis $\tilde{k}$ as the degree to which the distribution $S$ exhibits `heavy-tail phenomena', or the enhancement of the probability of what should otherwise be rare events.  Given that all Gaussian distributions have $\tilde{k}=0$, all measurements are then made relative to this fixed level.  Moreover, following the approach in \cite{dysthe2,dysthe3}, by comparing the initial spectral density to that at later times, we can see directly whether MI manifests as significant distortions of the statistics, or if through suppression the distributions remain relatively close to one another.

>>>>

\section{Introduction}
There is now a wide range of literature which shows that nonlinear instabilities, in particular the modulational instability (MI), are responsible for significant modifications to the statistical properties of water waves; see \cite{dysthe2,dysthe3,onorato,thomas2012nonlinear} among others.   As explained in \cite{janssen}, MI is especially important due to the relative rapidity with which it acts, being orders of magnitude faster than say four-wave resonant interactions which form the traditional backbone of our understanding of how nonlinearity drives changes in statistics of random-wave fields \cite{hasselmann,holthuijsen}.  A detailed set of theoretical and experimental results \cite{sergeeva1,sergeeva2,sergeeva3,sergeeva4} have also shown that MI could play a significant role in unidirectional deep-water-rogue wave formation among other oceanographic phenomena.  It is important to note though that there are several possible mechanisms for rogue-wave generation; see \cite{fedele} for example which rules out MI as a mechanism altogether.

However, much of the understanding around MI relies on highly idealized assumptions which cannot be expected to hold in natural settings.  In particular, most analytic understandings of MI rely on looking at the stability of perturbations to carrier waves at essentially a fixed wave number.  To address this shortcoming, in now seminal papers, \cite{alber} and \cite{alber2} analytically studied families of perturbations of wave packets with narrow, but non-zero, spectral width around a central wave number.  Using the Nonlinear Schr\"{o}dinger equation (NLSE), an analytic criterion determining when MI is either manifested or suppressed depending on this spectral width, now referred to as a Benjamin--Feir index (BFI) was derived.  The results in \cite{alber} were explored numerically in \cite{dysthe2,dysthe3} by examining the mean properties associated with ensembles of initial conditions.

@@

Later work has simulated the impact of wind and damping on the interplay between the BFI and the kurtosis of a wave field; see \cite{slunyaev,eeltink}.  However, the impact of vorticity was not taken into account in these works, thereby ignoring a central mechanism for the transfer of energy across length scales in oceanic flows.  Moreover, as shown recently in \cite{curtis8}, linear shear profiles can have strong impacts on the behavior of mean properties of surface wave flows with some shear profiles being able to suppress the Stokes Drift at the surface.  Therefore while a relatively idealized case of shear, the interplay between nonlinearity and constant vorticity can manifest in markedly different physics.  Building then on the work in \cite{thomas2012nonlinear} and \cite{curtis8}, and complementing the results in \cite{touboul}, we numerically explore the statistical properties of random-nonlinear-wave fields moving over deep water and linear shear profiles.  Moving beyond just examining the properties of the NLSE with vorticity, we look at the statistical properties of solutions to a higher-order model, the vor-Dysthe equation (VDE) derived in \cite{curtis8}.  As shown in \cite{onorato,janssen} and \cite{eeltink}, the higher-order terms associated with the Dysthe equation can have significant impacts on the statistics of the waves, so it is a nontrivial question to examine how the vorticity interacts with these higher-order nonlinearities.  Moreover, as shown in \cite{sergeeva2}, the NLSE can rapidly lose any quantitative predictive power in experimental settings while the Dysthe equation provides far more accurate results over a relatively wide range of experimental conditions.

Using our simulations, starting from normally distributed random-wave fields, we show that the BFI condition with respect to predicting stability is not especially useful, conforming to the results in \cite{janssen}.  While we are able to show that for sufficiently wide enough initial distributions that the temporal modifications to the profile are slight, which echoes the zero-vorticity results in \cite{sergeeva1,sergeeva2,sergeeva3}, whether or not this corresponds to an actual bifurcation in stability or simply reflects that for a wide enough profile the sideband growth affiliated with the MI has little impact is not clear.  We note though that classic MI analysis does provide a useful prediction of the time scale over which the most significant dynamics in the statistics of the wave field occur.

@@

The structure of the paper is an explanation of the model and discussion of the BFI in Section 2.  Section 3 presents our numerical results, while Section 4 presents a conclusion.  Explorations of two significant technical points are given in the Appendix.

\section{Modeling and Stability Theory}
We examine the unsteady nonlinear wave propagation over a constant shear current.  To do this, we suppose the fluid velocity is of the form
\[
{\bf u} = u({\bf x},t)\hat{{\bf x}} + w({\bf x},t)\hat{{\bf z}} = \omega z\hat{{\bf x}} + \nabla \phi,
\]
where $\phi$ is a harmonic function.  We restrict fluid motion to the $(x,z)$-plane, thereby ignoring transverse variations in the $y$ dimension.   Using well established arguments, e.g.~see \cite{ashton}, the dynamics of the fluid are found by solving the free boundary value problem
\begin{align}
\Delta \phi = 0, & & -\infty< z < \eta(x,t), \nonumber \\
\eta_{t} +\left(\omega\eta+\phi_{x}\right)\eta_{x} - \phi_{z} = 0, & & z=\eta(x,t), \label{kine}\\
\phi_{t} + \omega\pd_{x}^{-1}\eta_{t} + \frac{1}{2} \left|{\bf
    u}\right|^{2} + g\eta -
\frac{\sigma}{\rho}\pd_{x}\frac{\eta_{x}}{\sqrt{1+\eta_{x}^{2}}}= 0, &
& z=\eta(x,t), \label{bernou}\\
\lim_{z\rightarrow -\infty} \phi_{z} = 0, \nonumber
\end{align}
where $\eta$ represents the free surface displacement, and $g$, $\rho$, and $\sigma$ represent the acceleration due to gravity, the fluid density, and the coefficient of surface tension respectively.  The system is non-dimensionalized via the choices 
\begin{align*}
\tilde{x} = \frac{x}{L_{p}}, ~\tilde{z} = \frac{z}{L_{p}}, ~ \tilde{t} = \sqrt{\frac{g}{L_{p}}}~t, ~\omega = \sqrt{\frac{g}{L_{p}}}~\tilde{\omega}, \\
\eta = a \tilde{\eta}, ~ \phi  = a\sqrt{gL_{p}}~\tilde{\phi} , ~ \epsilon = \frac{a}{L_{p}}.
\end{align*}
Note that we are assuming $\omega = \mathcal{O}(1)$ throughout the remainder of the paper.  In physical terms, this implies that $\omega$ is comparable to the natural time scale of this problem, $\sqrt{L_{p}/g}$.  Letting $L_{p}\sim 1~m$, then this corresponds to the horizontal velocity of the fluid $u \sim 300 ~cm/s$ at a depth of one meter below the fluid surface.  Thus we are modeling the impact of strong surface currents moving over deep water.  If we were to assume $\omega$ were of larger magnitude, the problem would no longer be weakly nonlinear and asymptotic analysis would be significantly more difficult.  Therefore, throughout the remainder of the paper, we assume that the vorticity is not too large.  We also note that as $z\rightarrow -\infty$, the horizontal fluid velocity becomes unbounded in this model, which is physically absurd.  While we are only interested in results near the surface in this work, in the Appendix, we explore this issue and show that a more physically realistic shear profile would produce similar results at the surface to those presented throughout the remainder of the paper.

For a relatively small free-fluid surface over infinitely-deep water $z=\epsilon \eta(x,t)$, we model a modulated-wave packet via the expansion 
\begin{multline*}
\eta(x,t) = \eta_{1}(\xi,\tau)e^{i\theta(x,t)} + \eta^{\ast}_{1}(\xi,\tau)e^{-i\theta(x,t)}\\
+ \epsilon\left(\eta_{0}(\xi,\tau) + \eta_{2}(\xi,\tau)e^{2i\theta(x,t)} + \eta^{\ast}_{2}(\xi,\tau)e^{-2i\theta(x,t)}\right) + \mathcal{O}\left(\epsilon^{2} \right)
\end{multline*}
where $\tau = \epsilon^{2}t$, $\xi = \epsilon(x+c_{g}t)$, and $\theta(x,t) = k_{0}x + \Omega(k_{0},\omega)t$, where the linear dispersion relationship $\Omega(k_{0},\omega)$ is given by
\[
\Omega(k_{0},\omega) = \frac{1}{2}\left(s\omega \pm \sqrt{\omega^{2} + 4|k_{0}|\left(1+\tilde{\sigma}_{s}k_{0}^{2}\right)} \right).
\]
We note that the form for $\phi$ follows directly from this expansion, allowing for everything to be modeled by the terms in the expansion in $\eta$.  Here, $s = \mbox{sgn}(k_{0})$, $\omega$ is the non-dimensionalized magnitude of the vorticity of the flow, and $\tilde{\sigma}_{s}$ is the surface tension, or where the reciprocal of the Bond number, $\tilde{\sigma}$, is given by
\[
\tilde{\sigma}_{s} = \frac{\sigma}{\rho g L_{p}^2}.
\]
Throughout the remainder of this paper, we take the positive branch $+$ of the dispersion relationship, thereby ensuring that $\Omega > 0$.  Thus, for $k_{0}>0$, the surface waves to leading order propagate to the left.  Thus, if $\omega>0$, then throughout the bulk of the fluid, the shear current propagates with the surface waves, or is {\it co-propagating}, while for $\omega<0$, we say the shear current is {\it counter-propagating}.

To describe the evolution of the slowly evolving envelope $\eta_{1}(\xi,\tau)$, we use two models.  The first model is the the nonlinear Schr\"{o}dinger equation (NLSE) derived in \cite{curtis8}, though see also \cite{thomas2012nonlinear}, given by 
\begin{equation}
i\pd_{\tau}\eta_{1} + \alpha_{d}\pd_{\xi}^{2}\eta_{1} + \alpha_{nl}\left|\eta_{1} \right|^{2}\eta_{1} = 0, 
\label{nlse}
\end{equation}
where 
\begin{align*}
c_{g}(k_{0},\omega) = & \frac{1+3\tilde{\sigma}_{s}k_{0}^{2}}{2s\Omega - \omega},\\
\alpha_{d}(k_{0},\omega) = & \frac{(c^2_{g} - 3|k_{0}|\tilde{\sigma}_{s})}{2\Omega-s\omega},\\
\alpha_{nl}(k_{0},\omega) = & \frac{k_{0}\left( sk_{0}^{3}\left(8 + \tilde{\sigma}_{s}k_{0}^{2} + 2(\tilde{\sigma}_{s}k_{0}^{2})^{2}\right) + \omega \alpha_{v}\right)}{\left(2s\Omega -\omega\right)(1+c_{g}\omega)\left(4\Omega^2-s(2k_{0}(1+4\tilde{\sigma}_{s}k_{0}^{2})+2\omega\Omega)\right)},\\
\alpha_{v}(k_{0},\omega) = & s(c_g k_0  - 2\Omega ) \omega^4 + k_{0}(4 k_{0}^{2} s \tilde{\sigma}_{s} + 2 \Omega c_g  -s) \omega^3.
\end{align*}
However, the NLSE is a relatively low-order model known to be only of limited use in real-world modeling situations; see \cite{sergeeva2} in particular.  A more robust model, following the work in \cite{dysthe1}, is found by going to the next higher order in $\epsilon$, giving us the equation
\begin{multline}
(\omega-2s\Omega)\pd_{\tau}\eta_{1} + i(c_{g}^{2}s-3k_{0}\tilde{\sigma}_{s})\pd_{\xi}^{2}\eta_{1} + 2i\epsilon s c_{g}\pd^{2}_{\xi\tau}\eta_{1} - \epsilon\tilde{\sigma}_{s} \pd^{3}_{\xi}\eta_{1} + i\epsilon^{2}c_{h}\pd_{\xi}^{4}\eta_{1} \\
- i \alpha_{nl}(\omega-2s\Omega)\left|\eta_{1}\right|^{2}\eta_{1}
+ \epsilon \tilde{\alpha}_{1}\left|\eta_{1}\right|^{2}\pd_{\xi}\eta_{1}\\
+ \epsilon\tilde{\alpha}_{2}\eta_{1}^{2}\pd_{\xi}\eta^{\ast}_{1} +
\epsilon i\tilde{\alpha}_{3}\eta_{1}\mathcal{H}\pd_{\xi}\left|\eta_{1}\right|^{2}= 0.
\label{dysthe}
\end{multline}
where $\mathcal{H}$ denotes the Hilbert transform so that 
\[
\mathcal{H}f = \frac{1}{2\pi}\int_{\mathbb{R}}dk~ e^{ikx}  i\mbox{sgn}(k) \hat{f}(k),
\]
and where $\hat{f}(k)$ is the Fourier transform of $f(x)$ defined by
\[
\hat{f}(k) = \int_{\mathbb{R}}dx~ e^{-ikx}f(x).
\]

We call Equation \eqref{dysthe} the Vor-Dysthe equation (VDE).  Details of its derivation can be found in \cite{curtis8}, though we note a difference in this work with that in \cite{curtis8}.  In this work, we found the inclusion of the higher-order dispersion associated with the coefficient $c_{h}$, given by
\[
c_{h} = 2sc_{g}(2sc_{g}\alpha_{d}+\tilde{\sigma}_{s})/(\omega-2s\Omega) - s\alpha_{d}^{2},
\]
was necessary to remove artificial instabilities.  Similar issues with Dysthe like equations have been noted in \cite{eeltink} among other sources; see the Appendix for a discussion of this issue.  Moreover, the coefficients $\tilde{\alpha}_{j}$ are readily derived, but their forms are cumbersome and thus we omit explicitly writing them for the sake of readability.

The traditional definition of modulational instability (MI) begins by studying the stability of the plane-wave, or Stokes wave, solutions of the NLS equation given by
\[
\eta_{1}(\xi,\tau) = A e^{i\alpha_{nl}A^{2}\tau},
\]
where $A>0$ is a real constant.  Complementing the results in \cite{thomas2012nonlinear}, we study the stability of these solutions by considering perturbed solutions of the form
\begin{equation}
\eta_{1p}(\xi,\tau) = \Big{(}A +\mu \big{(}u_{p}(\xi,\tau)+iv_{p}(\xi,\tau)\big{)} +\mathcal{O}(\mu^2)\Big{)}\mbox{e}^{i\alpha_{nl}A^{2}\tau},
\label{pert}
\end{equation}
where $\mu$ is a small real parameter and $u_{p}$ and $v_{p}$ are real-valued functions.  Substituting Equation \eqref{pert} into the NLSE given by Equation \eqref{nlse} and linearizing gives
\[
\pd_{\tau}\bp u_{p}\\ v_{p}\ep = \bp 0 & -\alpha_{d}\pd_{\xi}^{2} \\ \alpha_{d}\pd_{\xi}^{2} + 2A^{2}\alpha_{nl} & 0 \ep \bp u_{p} \\ v_{p} \ep.
\]
Separating variables, applying a Fourier transform in $\xi$, and introducing the $\xi$ wave number $l$, establishes that NLS plane-wave solutions are unstable with respect to MIs if
\[
0 < l^{2} \leq 2\frac{\alpha_{nl}}{\alpha_{d}}A^{2}.
\]
This reduces to the classic requirement that MIs are suppressed if the coefficients of dispersion and nonlinearity have opposite signs (i.e.~the defocusing case).  Thus in the Results section we make choices of $\omega$ which keep us in the focusing case.  See \cite{curtis8} for further details.

As noted though in \cite{alber}, this notion of MI is only defined for perturbations around a monochromatic wave train.  We are thus motivated to examine perturbations around more complicated wavetrains.  To generate these simulations, we start both the NLSE and VDE from the same collection of $2L$ periodic-initial conditions
\[
\eta_{1}(\xi,0) = \epsilon_{rms}\sqrt{\frac{\delta \tilde{k}}{\sigma\sqrt{\pi}}} \sum_{k=-K+1}^{K}e^{-\tilde{k}^{2}/2\sigma^{2}}e^{i\theta_{k}}e^{i\tilde{k}\xi}, ~ \tilde{k} = \frac{\pi k}{L},
\]
where $\delta \tilde{k}=\pi/L$ and the phases $\theta_{k}$ are randomly and independently chosen uniformly between $0$ and $2\pi$.  We can then readily show that, for $L\gg 1$, 
\[
\epsilon_{rms} \approx \overline{\left|\eta_{1}(\xi,0)\right|^{2}}^{1/2},
\]
where $\overline{()}$ denotes the ensemble average.  In our coordinates then, motivated by the work of \cite{alber} and \cite{dysthe2} and more fully explored in \cite{janssen}, we study the stability of the initial wavepacket distributions with regards to the magnitude of the Benjamin--Feir Index (BFI), where the BFI is defined to be  
\[
\mbox{BFI} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\frac{\epsilon_{rms}}{\sigma}.
\]
Note, this definition is identical with the spectral width criterion of \cite{alber} and the BFI condition given in \cite{thomas2012nonlinear}; see also \cite{onorato}.  It was claimed in \cite{alber}, and in some sense verified in \cite{dysthe2}, that wave packets should be stable with respect to MIs for $\mbox{BFI}<1$, which corresponds to the initial spectral width $\sigma>\sigma_{c}$, where 
\[
\sigma_{c} = \sqrt{\frac{\alpha_{nl}}{\alpha_{d}}}\epsilon_{rms}.
\]
